#  Sample size investigation - Continued {#model-suiteC}

Repeat the analyses from Section \@ref(model-suite), but using
shuffled labels for the fitting set.  



## Run simulations - shuffled label lasso


To shuffle the labels we can simply swap rows in `sim_stage_lcpm_mtx`
between affected and control samples for each simulation.


<!-- RUN ONCE - THEN GET FROM MEMORY -->
```{r shuffled-run-sim, cache=T, cache.vars='start_time', eval=T}
### CLEAR CACHE
start_time <- proc.time()

# Get stage from SIZE
stage_vec <- cut(1:nrow(sim_control_mtx), c(0, SIZE), include.lowest = T)

# ran in two runs 1:7, 8:ncol
for (SIMno in 1:ncol(sim_control_mtx)) {

  cat("Running simulation ", SIMno, "\n")

  sim_cv_lst <- lapply(1:length(levels(stage_vec)), function(STGno) {
    Stage_rows_vec <- which(stage_vec %in% levels(stage_vec)[1:STGno])
    #cat("Stage ", STGno, "- analyzing", length(Stage_rows_vec), "paired samples.\n")

    sim_stage_samples_vec <- c(
      all_control_vec[sim_control_mtx[Stage_rows_vec, SIMno]],
      all_affected_vec[sim_affected_mtx[Stage_rows_vec, SIMno]]
    )

    sim_stage_lcpm_mtx <- all_lcpm_mtx[sim_stage_samples_vec, ]
    sim_stage_group_vec <- all_group_vec[sim_stage_samples_vec]

    Control_rows <- which(sim_stage_group_vec == 'Control')
    Affected_rows <- which(sim_stage_group_vec == 'HCC')

    swap_Control_ndx <- Control_rows[1:round(length(Control_rows)/2)]
    swap_Affected_ndx <- Affected_rows[1:round(length(Affected_rows)/2)]

    swap_Control_lcpm_mtx <- sim_stage_lcpm_mtx[swap_Control_ndx,]
    swap_Affected_lcpm_mtx <- sim_stage_lcpm_mtx[swap_Affected_ndx,]

    swap_stage_lcpm_mtx <- sim_stage_lcpm_mtx
    swap_stage_lcpm_mtx[swap_Control_ndx, ]  <- swap_Affected_lcpm_mtx
    swap_stage_lcpm_mtx[swap_Affected_ndx, ] <- swap_Control_lcpm_mtx
   
    #print(table(sim_stage_group_vec))

    sim_stage_cv_lst <- lapply(1:CV_REP, function(CV) {
      cv_fit <- glmnet::cv.glmnet(
        x = swap_stage_lcpm_mtx,
        y = sim_stage_group_vec,
        alpha = 1,
        family = "binomial",
        type.measure = "class",
        keep = T,
        nlambda = 30
      )

      # Extract 1se metrics from cv_fit
      #######################
      ndx_1se <- which(cv_fit$lambda == cv_fit$lambda.1se)

      nzero_1se <- cv_fit$nzero[ndx_1se]
      cvm_1se <- cv_fit$cvm[ndx_1se]

      # test error
      sim_stage_test_samples_vec <- setdiff(rownames(all_lcpm_mtx), sim_stage_samples_vec)
      sim_stage_test_lcpm_mtx <- all_lcpm_mtx[sim_stage_test_samples_vec,]
      sim_stage_test_group_vec <- all_group_vec[sim_stage_test_samples_vec]

      test_pred_1se_vec <- predict(
       cv_fit,
       newx=sim_stage_test_lcpm_mtx,
       s="lambda.1se",
       type="class"
      )
      test_1se_error <- mean(test_pred_1se_vec != sim_stage_test_group_vec)

      # genes
      coef_1se <- coef(
        cv_fit,
        s = "lambda.1se"
      )
      genes_1se <- coef_1se@Dimnames[[1]][coef_1se@i[-1]]

      # Extract min metrics from cv_fit
      #######################
      ndx_min <- which(cv_fit$lambda == cv_fit$lambda.min)

      nzero_min <- cv_fit$nzero[ndx_min]
      cvm_min <- cv_fit$cvm[ndx_min]

      # test error
      sim_stage_test_samples_vec <- setdiff(rownames(all_lcpm_mtx), sim_stage_samples_vec)
      sim_stage_test_lcpm_mtx <- all_lcpm_mtx[sim_stage_test_samples_vec,]
      sim_stage_test_group_vec <- all_group_vec[sim_stage_test_samples_vec]

      test_pred_min_vec <- predict(
       cv_fit,
       newx=sim_stage_test_lcpm_mtx,
       s="lambda.min",
       type="class"
      )
      test_min_error <- mean(test_pred_min_vec != sim_stage_test_group_vec)

      # genes
      coef_min <- coef(
        cv_fit,
        s = "lambda.min"
      )
      genes_min <- coef_min@Dimnames[[1]][coef_min@i[-1]]

      # return cv_fit summary metrics
      list(
       p_1se = nzero_1se, 
       p_min = nzero_min, 
       cv_1se = cvm_1se, 
       cv_min = cvm_min, 
       test_1se=test_1se_error, 
       test_min=test_min_error, 
       genes_1se = genes_1se,
       genes_min = genes_min)
    })
    sim_stage_cv_lst
  })

  # save  sim_cv_lst
  fName <- paste0("shuffledLasso_sim_", SIMno, "_cv_lst")
  assign(fName, sim_cv_lst)
  save(list = fName, file=file.path("RData", fName))

}
  message("simulation time: ", round((proc.time() - start_time)[3], 2), "s")

```


## shuffledLasso Simulation results

<!--
First we extract simluation results and store into one big table:
-->
```{r shuffledLasso-extract-sim-results, cache=T, cache.vars='shuffledLasso-run-sim', include=F}
# CLEAR CACHE 

sim_files_vec <- list.files('RData', '^shuffledLasso_sim')


# define extraction methods

# Each sumulation is a list of cv results 
## nested in a list of replicates
##############################################

# cvList2frm_f makes a frame out of the inner list
cvList2frm_f <- function(cv_lst) {
 frm1 <- as.data.frame(t(sapply(cv_lst, function(x) x)))
 frm2 <- data.frame(
  unlist(frm1[[1]]), unlist(frm1[[2]]),
  unlist(frm1[[3]]), unlist(frm1[[4]]),
  unlist(frm1[[5]]), unlist(frm1[[6]]),
  frm1[7], frm1[8])
  names(frm2) <- names(frm1)
  data.frame(Rep=1:nrow(frm2), frm2)}

# cv_lst_to_frm loop over replicates, concatenating the inner list frames
cv_lst_to_frm <- function(sim_cv_lst) {
 do.call('rbind', lapply(1:length(sim_cv_lst),
  function(JJ) {
    siz_frm <- cvList2frm_f(sim_cv_lst[[JJ]])
    data.frame(Size=SIZE[JJ], siz_frm)
  }))
}

# we loop across simulations to combine all results into one big table
shuffledLasso_sim_results_frm <- do.call('rbind', lapply(1:length(sim_files_vec),
 function(SIM_NO) {
  load(file=file.path('RData', sim_files_vec[SIM_NO]))
  assign('sim_cv_lst', get(sim_files_vec[SIM_NO]))
  rm(list=sim_files_vec[SIM_NO])
  
  data.frame(SimNo=paste0('Sim_',formatC(SIM_NO,width = 2,flag = 0)), cv_lst_to_frm(sim_cv_lst))
} 
)) 

```

<!-- 
Have a table of simulation results - `shuffledLasso_sim_results_frm`:
-->

```{r shuffledLasso-sum-table, cache=T, cache.vars='', fig.cap='Simution results table', include=F}

# CLEAR CACHE 
 
knitr::kable(head(with(shuffledLasso_sim_results_frm, table(SimNo, Size))),
  caption = paste("Simulation Results - N Sim =", SIM)) %>%
   kableExtra::kable_styling(full_width = F)

knitr::kable(head(shuffledLasso_sim_results_frm) %>% dplyr::select(-c(genes_1se, genes_min)),
    caption = paste("Simulation Results - not showing genes column"),
    digits=2) %>%
   kableExtra::kable_styling(full_width = F)

```


### shuffledLasso Simulation Results Look at one simulation


First examine results for one simulation run.

```{r shuffledLasso-simRes-errors-bySim, cache=T, cache.vars='what-to-keep', fig.heigth=5, fig.width=10, fig.cap='shuffledLasso Model Errors by Sample Size'}
### CLEAR CACHE

# get full model cv error ref
error_1se_vec <- sapply(cv_shuffledLassoAll_lst,
 function(cv_fit) cv_fit$cvm[cv_fit$lambda == cv_fit$lambda.1se])
error_1se_q2 <- quantile(error_1se_vec, prob=1/2)        

error_min_vec <- sapply(cv_shuffledLassoAll_lst,
 function(cv_fit) cv_fit$cvm[cv_fit$lambda == cv_fit$lambda.min])
error_min_q2 <- quantile(error_min_vec, prob=1/2)        

# Utility objects
SIZE0 <- stringr::str_pad(SIZE, width=3, pad='0')
stage_vec <- cut(1:nrow(sim_control_mtx), c(0,SIZE), include.lowest = T)


#SIM <- "Sim_01"

for(SIM in unique(shuffledLasso_sim_results_frm$SimNo)[1]){

SimNum <- as.numeric(sub('Sim_','',SIM))

simNo_results_frm <- shuffledLasso_sim_results_frm %>% dplyr::filter(SimNo==SIM)


# errors
par(mfrow=c(1,2), mar=c(4, 2, 2, 1), oma=c(0,0,2,0))
###################
# 1se
####################
cv_1se_lst <- with(simNo_results_frm,
 split(cv_1se, Size))
names(cv_1se_lst) <- paste0(stringr::str_pad(names(cv_1se_lst), width=3, pad='0'),'_cv')

test_1se_lst <- with(simNo_results_frm,
 split(test_1se, Size))
names(test_1se_lst) <- paste0(stringr::str_pad(names(test_1se_lst), width=3, pad='0'),'_cv')

error_1se_lst <- c(cv_1se_lst, test_1se_lst)
error_1se_lst <- error_1se_lst[order(names(error_1se_lst))]

boxplot(error_1se_lst, 
  border=c('blue','green'), 
  ylim=c(0.05, .4),
  xaxt='n'
)
LL <- -1
axis(side=1, tick=F, line = LL,
  at = match(paste0(SIZE0,'_cv'),names(error_1se_lst)), 
  SIZE0
 )
abline(v= match(paste0(SIZE0,'_cv'),names(error_1se_lst))[-1] - 0.5, col='grey')
abline(h= error_1se_q2, col = 'red')
legend('topright', 
   #title='1se errors', title.col = 'black',
   text.col = c('blue','green'),
   legend = c('cv error', 'test set'),
   bty='n'
 )
title(paste('one se lambda - error rates'))

SKIP  <- function() {
# Add qual annotation
control_qual_vec <- sapply(split(sim_control_qual_mtx[,SimNum], stage_vec), median)
affected_qual_vec <- sapply(split(sim_affected_qual_mtx[,SimNum], stage_vec), median)
LL <- LL + 1
axis(side=1, tick=F, line = LL,
  at = match(paste0(SIZE0,'_cv'),names(error_1se_lst)),
  round(control_qual_vec, 2)
 )
LL <- LL + 1
axis(side=1, tick=F, line = LL,
  at = match(paste0(SIZE0,'_cv'),names(error_1se_lst)),
  round(affected_qual_vec, 2)
 )
}#SKIP

# min
####################
cv_min_lst <- with(simNo_results_frm,
 split(cv_min, Size))
names(cv_min_lst) <- paste0(stringr::str_pad(names(cv_min_lst), width=3, pad='0'),'_cv')

test_min_lst <- with(simNo_results_frm,
 split(test_min, Size))
names(test_min_lst) <- paste0(stringr::str_pad(names(test_min_lst), width=3, pad='0'),'_cv')

error_min_lst <- c(cv_min_lst, test_min_lst)
error_min_lst <- error_min_lst[order(names(error_min_lst))]

boxplot(error_min_lst, 
  border=c('blue','green'), 
  ylim=c(0.05, .4),
  xaxt='n'
)
LL <- -1
axis(side=1, tick=F, line = LL,
  at = match(paste0(SIZE0,'_cv'),names(error_min_lst)), 
  SIZE0
 )
abline(v= match(paste0(SIZE0,'_cv'),names(error_min_lst))[-1] - 0.5, col='grey')
abline(h= error_min_q2, col = 'red')
legend('topright', 
   #title='min errors', title.col = 'black',
   text.col = c('blue','green'),
   legend = c('cv error', 'test set'),
   bty='n'
 )
title(paste('min lambda - error rates'))

SKIP  <- function() {
# Add qual annotation
control_qual_vec <- sapply(split(sim_control_qual_mtx[,SimNum], stage_vec), median)
affected_qual_vec <- sapply(split(sim_affected_qual_mtx[,SimNum], stage_vec), median)
LL <- LL + 1
axis(side=1, tick=F, line = LL,
  at = match(paste0(SIZE0,'_cv'),names(error_min_lst)),
  round(control_qual_vec, 2)
 )
LL <- LL + 1
axis(side=1, tick=F, line = LL,
  at = match(paste0(SIZE0,'_cv'),names(error_min_lst)),
  round(affected_qual_vec, 2)
 )
}#SKIP
mtext(side=3, outer=T, cex=1.25, paste('Sim =',  SIM))

} # for(SIM

```


```{r shuffledLasso-simRes-features-bySim, cache=T, cache.vars='what-to-keep', fig.heigth=5, fig.width=10, fig.cap='shuffledLasso Models Selected Features by Sample Size'}
### CLEAR CACHE

# get full model nzero ref
nzero_1se_vec <- sapply(cv_shuffledLassoAll_lst,
 function(cv_fit) cv_fit$nzero[cv_fit$lambda == cv_fit$lambda.1se])
nzero_1se_q2 <- quantile(nzero_1se_vec, prob=c(2)/4)

nzero_min_vec <- sapply(cv_shuffledLassoAll_lst,
 function(cv_fit) cv_fit$nzero[cv_fit$lambda == cv_fit$lambda.min])
nzero_min_q2 <- quantile(nzero_min_vec, prob=c(2)/4)

# Utility objects
SIZE0 <- stringr::str_pad(SIZE, width=3, pad='0')
stage_vec <- cut(1:nrow(sim_control_qual_mtx), c(0,SIZE), include.lowest = T)


#SIM <- "Sim_01"

for(SIM in unique(shuffledLasso_sim_results_frm$SimNo)[1]){

SimNum <- as.numeric(sub('Sim_','',SIM))

simNo_results_frm <- shuffledLasso_sim_results_frm %>% dplyr::filter(SimNo==SIM)


par(mfrow=c(1,2), mar=c(4, 2, 2, 1), oma=c(0,0,2,0))
###################
# 1se
####################
# selected feature counts
p_1se_lst <- with(simNo_results_frm,
 split(p_1se, Size))
names(p_1se_lst) <- paste0(stringr::str_pad(names(p_1se_lst), width=3, pad='0'),'_p')

# get selected features that are part of shuffledLasso_gene_sign_1se_vec
# - the signature selected genes
sign_genes_1se_lst <- lapply(1:nrow(simNo_results_frm), function(RR)
    intersect(unlist(simNo_results_frm[RR, 'genes_1se']), shuffledLasso_gene_sign_1se_vec))

sign_p_1se_lst <- split(sapply(sign_genes_1se_lst, length), simNo_results_frm$Size)
names(sign_p_1se_lst) <- paste0(stringr::str_pad(names(sign_p_1se_lst), width=3, pad='0'),'_signP')


p_singP_1se_lst <- c(p_1se_lst, sign_p_1se_lst)
p_singP_1se_lst <- p_singP_1se_lst[order(names(p_singP_1se_lst))]

boxplot(p_singP_1se_lst,
  border=c('blue','green'),
  #ylim=c(0, 300),
  xaxt='n'
)
LL <- -1
axis(side=1, tick=F, line = LL,
  at = match(paste0(SIZE0,'_p'),names(p_singP_1se_lst)),
  SIZE0
 )
abline(v= match(paste0(SIZE0,'_p'),names(p_singP_1se_lst))[-1] - 0.5, col='grey')
#abline(h= nzero_1se_q2, col = 'red')
legend('topleft',
   #title='1se errors', title.col = 'black',
   text.col = c('blue', 'green'),
   legend= c('selected genes','signature genes'),
   bty='n'
 )
title(paste('one se lamdba - selected gene counts'))

SKIP  <- function() {
# Add qual annotation
control_qual_vec <- sapply(split(sim_control_qual_mtx[,SimNum], stage_vec), median)
affected_qual_vec <- sapply(split(sim_affected_qual_mtx[,SimNum], stage_vec), median)
LL <- LL + 1
axis(side=1, tick=F, line = LL,
  at =  match(paste0(SIZE0,'_p'),names(p_singP_1se_lst)),
  round(control_qual_vec, 2)
 )
LL <- LL + 1
axis(side=1, tick=F, line = LL,
  at =  match(paste0(SIZE0,'_p'),names(p_singP_1se_lst)),
  round(affected_qual_vec, 2)
 )
}#SKIP

###################
# min
####################
# selected feature counts
p_min_lst <- with(simNo_results_frm,
 split(p_min, Size))
names(p_min_lst) <- paste0(stringr::str_pad(names(p_min_lst), width=3, pad='0'),'_p')

# get selected features that are part of shuffledLasso_gene_sign_min_vec
# - the signature selected genes
sign_genes_min_lst <- lapply(1:nrow(simNo_results_frm), function(RR)
    intersect(unlist(simNo_results_frm[RR, 'genes_min']), shuffledLasso_gene_sign_min_vec))

sign_p_min_lst <- split(sapply(sign_genes_min_lst, length), simNo_results_frm$Size)
names(sign_p_min_lst) <- paste0(stringr::str_pad(names(sign_p_min_lst), width=3, pad='0'),'_signP')


p_singP_min_lst <- c(p_min_lst, sign_p_min_lst)
p_singP_min_lst <- p_singP_min_lst[order(names(p_singP_min_lst))]

boxplot(p_singP_min_lst,
  border=c('blue','green'),
  #ylim=c(0, 300),
  xaxt='n'
)
LL <- -1
axis(side=1, tick=F, line = LL,
  at = match(paste0(SIZE0,'_p'),names(p_singP_min_lst)),
  SIZE0
 )
abline(v= match(paste0(SIZE0,'_p'),names(p_singP_min_lst))[-1] - 0.5, col='grey')
#abline(h= nzero_min_q2, col = 'red')
legend('topleft',
   #title='min errors', title.col = 'black',
   text.col = c('blue', 'green'),
   legend= c('selected genes','signature genes'),
   bty='n'
 )
title(paste('min lambda - selected gene counts'))

SKIP  <- function() {
# Add qual annotation
control_qual_vec <- sapply(split(sim_control_qual_mtx[,SimNum], stage_vec), median)
affected_qual_vec <- sapply(split(sim_affected_qual_mtx[,SimNum], stage_vec), median)
LL <- LL + 1
axis(side=1, tick=F, line = LL,
  at =  match(paste0(SIZE0,'_p'),names(p_singP_min_lst)),
  round(control_qual_vec, 2)
 )
LL <- LL + 1
axis(side=1, tick=F, line = LL,
  at =  match(paste0(SIZE0,'_p'),names(p_singP_min_lst)),
  round(affected_qual_vec, 2)
 )
}#SKIP

mtext(side=3, outer=T, cex=1.25, paste('Sim =',  SIM))

} # for(SIM

```


### Summarize results across simulation runs.

```{r shuffledLasso-simRes-errors-overSim, cache=T, cache.vars=c('error_1se_Bysize_lst','error_min_Bysize_lst'), fig.heigth=5, fig.width=10, fig.cap='shuffledLasso Model Errors by Sample Size'}
### CLEAR CACHE

# get full model cv error ref
error_1se_vec <- sapply(cv_shuffledLassoAll_lst,
 function(cv_fit) cv_fit$cvm[cv_fit$lambda == cv_fit$lambda.1se])
error_1se_q2 <- quantile(error_1se_vec, prob=1/2)        

error_min_vec <- sapply(cv_shuffledLassoAll_lst,
 function(cv_fit) cv_fit$cvm[cv_fit$lambda == cv_fit$lambda.min])
error_min_q2 <- quantile(error_min_vec, prob=1/2)        

# Utility objects
SIZE0 <- stringr::str_pad(SIZE, width=3, pad='0')
stage_vec <- cut(1:nrow(sim_control_qual_mtx), c(0,SIZE), include.lowest = T)

# 1se
#########################################
## cv
cv_1se_Bysize_lst <- lapply(unique(shuffledLasso_sim_results_frm$Size),
function(SizeVal) {
 sizeVal_results_frm <- shuffledLasso_sim_results_frm %>% dplyr::filter(Size==SizeVal)
 sizeVal_cv_1se_lst <- with(sizeVal_results_frm, split(cv_1se, SimNo))
 sapply(sizeVal_cv_1se_lst, median)
})
names(cv_1se_Bysize_lst) <- paste0(
 stringr::str_pad(unique(shuffledLasso_sim_results_frm$Size), width=3, pad='0'), '_cv')

## test
test_1se_Bysize_lst <- lapply(unique(shuffledLasso_sim_results_frm$Size),
function(SizeVal) {
 sizeVal_results_frm <- shuffledLasso_sim_results_frm %>% dplyr::filter(Size==SizeVal)
 sizeVal_test_1se_lst <- with(sizeVal_results_frm, split(test_1se, SimNo))
 sapply(sizeVal_test_1se_lst, median)
})
names(test_1se_Bysize_lst) <- paste0(
 stringr::str_pad(unique(shuffledLasso_sim_results_frm$Size), width=3, pad='0'), '_test')


error_1se_Bysize_lst <- c(cv_1se_Bysize_lst, test_1se_Bysize_lst)
error_1se_Bysize_lst <- error_1se_Bysize_lst[order(names(error_1se_Bysize_lst))]

boxplot(error_1se_Bysize_lst,
  col=0,
  border=c('blue','green'),
  ylim=c(0.05, .5),
  outline=F,
  xaxt='n'
)
for(JJ in 1:length(error_1se_Bysize_lst))
points(
   x=jitter(rep(JJ, length(error_1se_Bysize_lst[[JJ]])), amount=0.25), 
   y=error_1se_Bysize_lst[[JJ]],
   col=ifelse(grepl('cv', names(error_1se_Bysize_lst)[JJ]),'blue', 'green')
)
LL <- -1
axis(side=1, tick=F, line = LL,
  at = match(paste0(SIZE0,'_cv'),names(error_1se_Bysize_lst)),
  SIZE0
 )
abline(v= match(paste0(SIZE0,'_cv'),names(error_1se_Bysize_lst))[-1] - 0.5, col='grey')
abline(h= error_min_q2, col = 'red')
legend('topright',
   #title='min errors', title.col = 'black',
   text.col = c('blue','green'),
   legend = c('cv error', 'test set'),
   bty='n'
 )
title(paste('one se lambda - error rates'))


# min
#########################################
## cv
cv_min_Bysize_lst <- lapply(unique(shuffledLasso_sim_results_frm$Size),
function(SizeVal) {
 sizeVal_results_frm <- shuffledLasso_sim_results_frm %>% dplyr::filter(Size==SizeVal)
 sizeVal_cv_min_lst <- with(sizeVal_results_frm, split(cv_min, SimNo))
 sapply(sizeVal_cv_min_lst, median)
})
names(cv_min_Bysize_lst) <- paste0(
 stringr::str_pad(unique(shuffledLasso_sim_results_frm$Size), width=3, pad='0'), '_cv')

## test
test_min_Bysize_lst <- lapply(unique(shuffledLasso_sim_results_frm$Size),
function(SizeVal) {
 sizeVal_results_frm <- shuffledLasso_sim_results_frm %>% dplyr::filter(Size==SizeVal)
 sizeVal_test_min_lst <- with(sizeVal_results_frm, split(test_min, SimNo))
 sapply(sizeVal_test_min_lst, median)
})
names(test_min_Bysize_lst) <- paste0(
 stringr::str_pad(unique(shuffledLasso_sim_results_frm$Size), width=3, pad='0'), '_test')


error_min_Bysize_lst <- c(cv_min_Bysize_lst, test_min_Bysize_lst)
error_min_Bysize_lst <- error_min_Bysize_lst[order(names(error_min_Bysize_lst))]

boxplot(error_min_Bysize_lst,
  col=0,
  border=c('blue','green'),
  ylim=c(0.05, .5),
  outline=F,
  xaxt='n'
)
for(JJ in 1:length(error_min_Bysize_lst))
points(
   x=jitter(rep(JJ, length(error_min_Bysize_lst[[JJ]])), amount=0.25), 
   y=error_min_Bysize_lst[[JJ]],
   col=ifelse(grepl('cv', names(error_min_Bysize_lst)[JJ]),'blue', 'green')
)
LL <- -1
axis(side=1, tick=F, line = LL,
  at = match(paste0(SIZE0,'_cv'),names(error_min_Bysize_lst)),
  SIZE0
 )
abline(v= match(paste0(SIZE0,'_cv'),names(error_min_Bysize_lst))[-1] - 0.5, col='grey')
abline(h= error_min_q2, col = 'red')
legend('topright',
   #title='min errors', title.col = 'black',
   text.col = c('blue','green'),
   legend = c('cv error', 'test set'),
   bty='n'
 )
title(paste('min lambda - error rates'))


mtext(side=3, outer=T, cex=1.25, paste('shuffledLasso fit error rates summarized across simulations'))

```

```{r print-shuffledLasso-simRes-errors-overSim, cache=T, cache.vars='', fig.cap='shuffledLasso Model Errors by Sample Size'}
### CLEAR CACHE

error_1se_Bysize_sum_frm <- t(sapply(error_1se_Bysize_lst, function(LL) quantile(LL, prob=(1:3)/4)))

knitr::kable(error_1se_Bysize_sum_frm,
    caption = paste("1se lambda shuffledLasso error rates by sample size across simulations"),
    digits=2) %>%
   kableExtra::kable_styling(full_width = F)

error_min_Bysize_sum_frm <- t(sapply(error_min_Bysize_lst, function(LL) quantile(LL, prob=(1:3)/4)))

knitr::kable(error_min_Bysize_sum_frm,
    caption = paste("min lambda shuffledLasso error rates by sample size across simulations"),
    digits=2) %>%
   kableExtra::kable_styling(full_width = F)

```


Now look at feature selection.

```{r shuffledLasso-simRes-features-OverSim, cache=T, cache.vars=c('p_singP_1se_Bysize_lst','p_singP_min_Bysize_lst'), fig.heigth=5, fig.width=10, fig.cap='shuffledLasso Models Selected Features by Sample Size'}
### CLEAR CACHE

# Utility objects
SIZE0 <- stringr::str_pad(SIZE, width=3, pad='0')
stage_vec <- cut(1:nrow(sim_control_qual_mtx), c(0,SIZE), include.lowest = T)

# 1se
#########################################
# selected features
p_1se_Bysize_lst <- lapply(unique(shuffledLasso_sim_results_frm$Size),
function(SizeVal) {
 sizeVal_results_frm <- shuffledLasso_sim_results_frm %>% dplyr::filter(Size==SizeVal)
 sizeVal_p_1se_lst <- with(sizeVal_results_frm, split(p_1se, SimNo))
 sapply(sizeVal_p_1se_lst, median)
})
names(p_1se_Bysize_lst) <- paste0(
 stringr::str_pad(unique(shuffledLasso_sim_results_frm$Size), width=3, pad='0'), '_p')

# selected signatue features
sign_p_1se_Bysize_lst <- lapply(unique(shuffledLasso_sim_results_frm$Size),
function(SizeVal) {
 sizeVal_results_frm <- shuffledLasso_sim_results_frm %>% dplyr::filter(Size==SizeVal)
 
  
 sizeVal_sign_genes_1se_lst <- lapply(1:nrow(sizeVal_results_frm), function(RR)
    intersect(unlist(sizeVal_results_frm[RR, 'genes_1se']), shuffledLasso_gene_sign_1se_vec))

 sizeVal_sign_p_1se_lst <- split(sapply(sizeVal_sign_genes_1se_lst, length),
    sizeVal_results_frm$SimNo)
 
 sapply(sizeVal_sign_p_1se_lst, median)
})
names(sign_p_1se_Bysize_lst) <- paste0(
 stringr::str_pad(unique(shuffledLasso_sim_results_frm$Size), width=3, pad='0'), '_signP')


p_singP_1se_Bysize_lst <- c(p_1se_Bysize_lst, sign_p_1se_Bysize_lst)
p_singP_1se_Bysize_lst <- p_singP_1se_Bysize_lst[order(names(p_singP_1se_Bysize_lst))]

boxplot(p_singP_1se_Bysize_lst,
  col=0,
  border=c('blue','green'),
  #ylim=c(0, 300),
  xaxt='n'
)
for(JJ in 1:length(p_singP_1se_Bysize_lst))
points(
   x=jitter(rep(JJ, length(p_singP_1se_Bysize_lst[[JJ]])), amount=0.25),
   y=p_singP_1se_Bysize_lst[[JJ]],
   col=ifelse(grepl('_p', names(p_singP_1se_Bysize_lst)[JJ]),'blue', 'green')
)

LL <- -1
axis(side=1, tick=F, line = LL,
  at = match(paste0(SIZE0,'_p'),names(p_singP_1se_Bysize_lst)),
  SIZE0
 )
abline(v= match(paste0(SIZE0,'_p'),names(p_singP_1se_Bysize_lst))[-1] - 0.5, col='grey')
#abline(h= nzero_1se_q2, col = 'red')
legend('topleft',
   #title='1se errors', title.col = 'black',
   text.col = c('blue', 'green'),
   legend= c('selected genes','signature genes'),
   bty='n'
 )
title(paste('one se lamdba - selected gene counts'))


# min
#########################################
# selected features
p_min_Bysize_lst <- lapply(unique(shuffledLasso_sim_results_frm$Size),
function(SizeVal) {
 sizeVal_results_frm <- shuffledLasso_sim_results_frm %>% dplyr::filter(Size==SizeVal)
 sizeVal_p_min_lst <- with(sizeVal_results_frm, split(p_min, SimNo))
 sapply(sizeVal_p_min_lst, median)
})
names(p_min_Bysize_lst) <- paste0(
 stringr::str_pad(unique(shuffledLasso_sim_results_frm$Size), width=3, pad='0'), '_p')

# selected signatue features
sign_p_min_Bysize_lst <- lapply(unique(shuffledLasso_sim_results_frm$Size),
function(SizeVal) {
 sizeVal_results_frm <- shuffledLasso_sim_results_frm %>% dplyr::filter(Size==SizeVal)
 
  
 sizeVal_sign_genes_min_lst <- lapply(1:nrow(sizeVal_results_frm), function(RR)
    intersect(unlist(sizeVal_results_frm[RR, 'genes_min']), shuffledLasso_gene_sign_min_vec))

 sizeVal_sign_p_min_lst <- split(sapply(sizeVal_sign_genes_min_lst, length),
    sizeVal_results_frm$SimNo)
 
 sapply(sizeVal_sign_p_min_lst, median)
})
names(sign_p_min_Bysize_lst) <- paste0(
 stringr::str_pad(unique(shuffledLasso_sim_results_frm$Size), width=3, pad='0'), '_signP')


p_singP_min_Bysize_lst <- c(p_min_Bysize_lst, sign_p_min_Bysize_lst)
p_singP_min_Bysize_lst <- p_singP_min_Bysize_lst[order(names(p_singP_min_Bysize_lst))]

boxplot(p_singP_min_Bysize_lst,
  col=0,
  border=c('blue','green'),
  #ylim=c(0, 300),
  xaxt='n'
)
for(JJ in 1:length(p_singP_min_Bysize_lst))
points(
   x=jitter(rep(JJ, length(p_singP_min_Bysize_lst[[JJ]])), amount=0.25),
   y=p_singP_min_Bysize_lst[[JJ]],
   col=ifelse(grepl('_p', names(p_singP_min_Bysize_lst)[JJ]),'blue', 'green')
)

LL <- -1
axis(side=1, tick=F, line = LL,
  at = match(paste0(SIZE0,'_p'),names(p_singP_min_Bysize_lst)),
  SIZE0
 )
abline(v= match(paste0(SIZE0,'_p'),names(p_singP_min_Bysize_lst))[-1] - 0.5, col='grey')
#abline(h= nzero_min_q2, col = 'red')
legend('topleft',
   #title='min errors', title.col = 'black',
   text.col = c('blue', 'green'),
   legend= c('selected genes','signature genes'),
   bty='n'
 )
title(paste('min lambda - selected gene counts'))

mtext(side=3, outer=T, cex=1.25, paste('shuffledLasso fit feature selection summarized across simulations'))


```


```{r print-shuffledLasso-simRes-features-OverSim, cache=T, cache.vars='', fig.cap='shuffledLasso Models Selected Features by Sample Size'}
### CLEAR CACHE

p_sing_1se_Bysize_sum_frm <- t(sapply(p_singP_1se_Bysize_lst, function(LL) quantile(LL, prob=(1:3)/4)))

knitr::kable(p_sing_1se_Bysize_sum_frm,
    caption = paste("1se lambda shuffledLasso feature selection sample size across simulations"),
    digits=2) %>%
   kableExtra::kable_styling(full_width = F)


p_sing_min_Bysize_sum_frm <- t(sapply(p_singP_min_Bysize_lst, function(LL) quantile(LL, prob=(1:3)/4)))

knitr::kable(p_sing_min_Bysize_sum_frm,
    caption = paste("min lambda shuffledLasso feature selection sample size across simulations"),
    digits=2) %>%
   kableExtra::kable_styling(full_width = F)


```

