# Variable importance  {#variable-importance}

Cai et al. [@Cai:2019aa] used the fequency of selection across
bootstrap replicates to select features from their final model.
We suspect that this simply selects features with large coefficients
which would correspond the the variable importance metric
used in the `caret` package.  

In this section we measure variable importance as the loss in performance when the variable is left out
of the model.  Depending on the correlation structure, it is quite possible
for a feature to have a large coefficient and be un-important, in our sense of the word.
It is also possible for a feature to be frequently selected in bootstrap samples, and
still not un-important.  In a sense, this measure of importance is really a measure
of `single importance`.  We should therefore also have a measure of `group importance`.

For this exercise, wr will use the min lambda error assessment as it has
lower variability across repeated cv fits.

```{r getCoefMtx, cache=T, cache.vars=c('lassoAll_coefVal_lse_mtx')}
### CLEAR CACHE


# min
lassoAll_coef_min_lst <- lapply(cv_lassoAll_lst, function(cv_fit){
 cv_fit_coef <- coef(
 cv_fit,
 s = "lambda.min"
 )
 cv_fit_coef@Dimnames[[1]][cv_fit_coef@i[-1]]
 })



# min
lassoAll_coefVal_min_lst <- lapply(cv_lassoAll_lst, function(cv_fit){
 cv_fit_coef <- coef(
 cv_fit,
 s = "lambda.min"
 )
 coef_vec <- cv_fit_coef@x[-1]
 names(coef_vec) <-  cv_fit_coef@Dimnames[[1]][cv_fit_coef@i[-1]]

 coef_vec
 })



# put feature occurence into matrix
lassoAll_coef_min_all <- Reduce(union, lassoAll_coef_min_lst)
lassoAll_coef_min_mtx <- sapply(lassoAll_coef_min_lst, 
  function(LL) is.element(lassoAll_coef_min_all, LL)
)
rownames(lassoAll_coef_min_mtx) <- lassoAll_coef_min_all

# put festures coefficients into matrix
lassoAll_coefVal_lse_mtx <- matrix(
 nrow=nrow(lassoAll_coef_min_mtx), 
 ncol=ncol(lassoAll_coef_min_mtx))
rownames(lassoAll_coefVal_lse_mtx) <- rownames(lassoAll_coef_min_mtx)

for(FF in rownames(lassoAll_coefVal_lse_mtx))
for(SIM in 1:ncol(lassoAll_coefVal_lse_mtx))
lassoAll_coefVal_lse_mtx[FF, SIM] <- lassoAll_coefVal_min_lst[[SIM]][FF]
lassoAll_coefVal_lse_mtx[is.na(lassoAll_coefVal_lse_mtx)] <- 0

```


```{r varOccurence, cache=F, cache.vars=c('variables-to-keep'), fig.cap="one-standard-error rule lasso feature selection"}

# order by selection frequencey
feature_freq_vec <- rowSums(abs(lassoAll_coefVal_lse_mtx) > 0)
table(feature_freq_vec)

feature_o <- order(feature_freq_vec)

boxplot(
   t(lassoAll_coefVal_lse_mtx)[,feature_o], 
   ylab='Coefficient', xlab='Frequency of selection',
   xaxt='n', 
   outline=F)
abline(h=0, col='red')
axis(
 side=1, 
 at=match(unique(feature_freq_vec), sort(feature_freq_vec)),
 unique(feature_freq_vec))
 
 #title("one-standard-error rule lasso feature selection")

```

As we suspected, features with large absolute regularized coefficients tend to
be selected more frequently across repreated cross validated fits.   So essntially,
the Cai et al. frequecny filter is essentially a size of coefficient filter.

We not turn to this other notion of variable importance: what is the increase in
error rate when each variable is excluded from the fit.  We will us cv error rates in
the one-standard-error rule lasso models for this assessment.  As this computation
will be pretty costly, we will select 30 features from different
selection frequency blocks

```{r selFreq, commnent=''}

selFreq_bin_vec <- cut(feature_freq_vec, c(0,10, 15, 20, 25, 29,30))
table(selFreq_bin_vec)

set.seed(1)
sel_features_lst <- lapply(split(feature_freq_vec, selFreq_bin_vec),
  function(FF) sample(FF, size=5))

cat("List of selected features to assess:\n")
sel_features_lst

```


For each selected feature, run 10 cv fits with 10 lambda values, and compute the 
min lambda error.


## IN PROGESS {-}


```{r leaveOneOut, cache=T, cache.vars='cv_lassoLOO_lst', eval=F}

set.seed(1)

start_time <-  proc.time()

cv_lassoLOO_lst  <- lapply(as.vector(sapply(sel_features_lst, names)),
 function(FF) {
  FF_ndx <- match(FF, colnames(all_lcpm_mtx))

  cv_lassoMFF_lst <- lapply(1:10, function(REP) {
  glmnet::cv.glmnet(
   x = all_lcpm_mtx[,-FF_ndx],
   y = factor(all_group_vec,levels = c('Control', 'HCC')),
   alpha = 1,
   family = 'binomial',
   type.measure  =  "class",
   keep = F,
   nlambda = 10
   )
  })
 
  cbind(
 `1se` = sapply(cv_lassoMFF_lst, function(LL) 
          LL$cvm[LL$lambda==LL$lambda.1se]),
  `min` = sapply(cv_lassoMFF_lst, function(LL)
          LL$cvm[LL$lambda==LL$lambda.1se])
  )
  })

   save(list='cv_lassoMFF_lst', file=file.path("RData",'cv_lassoMFF_lst'))

```


